---
title: 'R tutorial for downloading Statsbomb event data'
subtitle: 'All Seasons'
author: 'Pablo L. Landeros'
date: "`r Sys.Date()`"
abstract: The aim of this writeup is to provide a tutorial on how to download data from the `StatsBombR` library
title-block-banner: false
execute:
  error: true
  warning: false
  message: false
toc-title: 'Content'
backgroundcolor: "#eee9e5"
format:
  html:
    theme: sandstone
    toc: true
    toc-location: left
    toc-depth: 3
    code-fold: false
    grid:
      body-width: 1500px
reference-location: margin
citation-location: margin
link-citations: true
---
## Installing Rtools
`Rtools` is required to install the Statsbomb libraries. To install: 

1. Go to [this link](https://cran.r-project.org/bin/windows/Rtools/rtools42/rtools.html) and download the Rtools installer.
2. After downloading has completed, run the installer. Select the default options everywhere.

## Library Installation
To install the `StatsbombR` package, you’ll need to install the `devtools` package:
```{r, eval = FALSE, echo = TRUE}
install.packages('devtools')
install.packages('tidyverse')
install.packages('remotes')
remotes::install_version("SDMTools", "1.1-221")
```
Then, to install `StatsbombR` itself, run:
```{r, eval = FALSE, echo = TRUE}
devtools::install_github("statsbomb/StatsBombR")
```

Load `tidyverse` and Statsbomb libraries:
```{r, eval = FALSE, echo = TRUE}
library(tidyverse)
library(SDMTools)
library(StatsBombR)
```

## Downloading Women's FA Super League Data

#### Establish competition ID
`competition_id` for Women's FA Super league should be `37`.
```{r, eval = FALSE, echo = TRUE}
comp_id <- 37  #competition_id of the FA Women's super League
```

#### Get season IDs
The IDs for the 3 available seasons are `90`,`42` and `2`:
```{r, eval = FALSE, echo = TRUE}
season_ids <- c(90, 42, 4)
```

#### Get data
This grabs the competitions that are available to the user and filters it
down, using dplyr’s `filter` function, to just the 2020/21 FA Women’s Super
League season in this example.
```{r, eval = FALSE, echo = TRUE}
comps <- FreeCompetitions() %>%
  dplyr::filter(competition_id==comp_id & season_id %in% season_ids) 
```
#### Exctract all free matches
Pull matches for the desired competition:
```{r, eval = FALSE, echo = TRUE}
matches <- FreeMatches(comps)
```
#### Get events
Create a table of the free event data for the FAWSL seasons.
```{r,eval = FALSE, echo = TRUE}
sb_data <- StatsBombFreeEvents(MatchesDF = matches, Parallel = T)
```
This might take a while to run

Finally, we can use Statsbomb's `allclean` function to quickly exctract lots of relevant information such as x/y coordinates. 

More information can be found in the package info. Be sure to familiarise
yourself with the columns it creates using names(`nameofyourdfhere`).
```{r, eval = FALSE, echo = TRUE}
event_data <- allclean(sb_data)
```

## Official Statsbomb Documentation
For more detailed documentation on Statsbomb data, you can check out the full documentation:

- [Open Data Competitions v2.0.0.pdf](https://github.com/statsbomb/open-data/blob/master/doc/Open%20Data%20Competitions%20v2.0.0.pdf) - Covers the objects contained within the competitions information
(FreeCompetitions() ).
- [Open Data Matches v3.0.0.pdf](https://github.com/statsbomb/open-data/blob/master/doc/Open%20Data%20Matches%20v3.0.0.pdf) - Describes the match info download ( FreeMatches()).
- [Open Data Lineups v2.0.0.pdf](https://github.com/statsbomb/open-data/blob/master/doc/Open%20Data%20Lineups%20v2.0.0.pdf) - Describes the structure of the lineup info ( getlineupsFree()).
- [Open Data Events v4.0.0](https://github.com/statsbomb/open-data/blob/master/doc/Open%20Data%20Events%20v4.0.0.pdf) - Explains the meaning of the column names within the event data.
- [StatsBomb Event Data Specification v1.1](https://github.com/statsbomb/open-data/blob/master/doc/StatsBomb%20Open%20Data%20Specification%20v1.1.pdf)- The full breakdown of all the events within the data.

Now you're ready to participate! ╭(◔ ◡ ◔)/

- [Home](https://pablolopez2733.github.io/Aztec-Data-Summit/)
- [Official StatsbombR documentation](https://statsbomb.com/wp-content/uploads/2021/11/Working-with-R.pdf)









